---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ECPIN

<!-- badges: start -->
<!-- badges: end -->

The current project aims to analyze protein-metabolite interaction data

## Installation

You can install the the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("huiUofT/ECPIN")
```
## Example

This is a basic example which shows you how to solve a common problem:

Step 1: loading packages
```{r, message=FALSE, warning=FALSE}
  rm(list=ls())
  library(xcms)
  library(EcoliPMI)
  library(devtools)
  library(xlsx)
  library(ChemmineR)
  library(rcdk)
  library(MassSpecWavelet)
  library(Rcpp)
  library(RcppArmadillo)
  library(isopat)
  library(readxl)
  library(ggplot2)
  library(tidyverse)
  library(dplyr)
  library(ggrepel)
  library(openxlsx)
  #'set up the path
  path<-getwd()
  polarity<-1
  setwd(path)
```

Step 2: calibrating mass spectra
```{r, include= FALSE}
  #setwd('C:/Rprogram/ECPIN')
  #MassCalibration(polarity)
```

Step 3: extracting peak features
```{r, message=FALSE, warning=FALSE}
  setwd(path)

  #extracting peaks, intensity cutoff is 10^6, ppm = 2.5
  peaks.raw<-Peakextract(10^5,2.5)
  
  #'Extracting differential peaks
  #'data has to be organized like NR_XX_X, WT_XX_X, NR_DMSO_X
  Control<-'MeOH'
  peaks<-LigandFeatureID(peaks.raw,'pos',Control)#This is the function extracting differential peaks compared to controL
```
Step 4: identifying primary isotopic peaks, and remove redudant isotopic peaks
```{r, message=FALSE, warning=FALSE}
  #2 ppm is the mass tolerance to find isotopic peaks
  #'0.80 is the correlation coefficient to extract isotopic peaks
  #'3 is the intensity ratio cutoff between primary and other isotopic peaks
  peaks.iso<-FindIsotope(peaks,2,0.80,2)
```
Step 5: loading database, and do initial searching with exact mass
```{r, message=FALSE, warning=FALSE}
  setwd(path)
  Database<-read_excel("HMDB_CFMID.xlsx")#This is a combined database of TSCA, Tox21 and HMDB
  peaks.iso$mz<-peaks.iso$mz*(1+0*10^(-6))##simple mass calibration
  Library<-InitialSearch(peaks.iso,3,polarity,Database)#2ppm mass tolerance
  
  #delete redundant peaks
  mz<-NULL
  rt<-NULL
  indexsave<-NULL
  for (i in 1:nrow(Library)){
    index<-which(mz==Library$mz[i])
    if (length(index)==0){#new mz
      mz<-c(mz,Library$mz[i])
      rt<-c(rt,Library$rt[i])
      next
    }
    if (Library$rt[i]==rt[index[1]]){#retention time is identical
      indexsave<-c(indexsave,i)
      next
    }
    mz<-c(mz,Library$mz[i])
    rt<-c(rt,Library$rt[i])
  }
  if (length(indexsave)>0){
  Library<-Library[-indexsave,]}
  
  #'writing the initial searching
  write.table(Library,file='Metabolites_STD.csv',sep=',',row.names = FALSE)
```
Step 6: calculate scores
```{r, message=FALSE, warning=FALSE}
  setwd(path)
  #' the path to save results
  path<-getwd()
  Library.score<-ScoreRT(Library,Database)
  output<-Output(Library.score,-1000)
  
  write.table(output,file='Metabolite_final.csv',sep=',',row.names = FALSE)
```
